---
title: "FishBae"
subtitle: "EEMB 595 Final Project" 
author: "Phoebe Racine & Erin Winslow"
date: "5/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Questions for Grace
1. Do we need to code in JAGS?
- we're considering just R and are considering STAN
2.The length of the presentation might be long considering where we're at vs. where the class is at. What might be more appropriate?

#Task List for Phoebe & Erin
- ~~copy/paste from Juanca~~
  - find CSV & data so code connects and runs
- copy/paste GLM code from R
- read 3.3 of stats book
- research alpha/beta for GLMs
- consider including RAM legacy data if we want to provide a fuller story
  - consider analysis type
- Presentation
  - pull graphs from early AquaForce
  



#Loading Packages
```{r}
library(tidyverse)


```


#Salmon and Lobster Code from Juanca

##Put data together
We are using FAO’s data for **aquaculture production** and **wild fisheries**.

```{r}
FAO_sp_groups <- read.csv(file = here("raw_data","FAO","FAO_spp_groups.csv"), stringsAsFactors = FALSE) %>% 
  janitor::clean_names() %>% 
  select(species = x3alpha_code, family = family_group, order = order_group) %>% 
  mutate(family = taxize::taxize_capwords(family, onlyfirst = T),
         order = taxize::taxize_capwords(order, onlyfirst = T))

# Load FAO species codes lookup table
FAO_sp_codes <- read.csv(file = here("raw_data","FAO","FAO_sp_codes.csv"), stringsAsFactors = FALSE) %>% 
  janitor::clean_names() %>% 
  select(species = x3alpha_code, common_name = name_en, sci_name = scientific_name) %>% 
  left_join(FAO_sp_groups, by = "species") %>% 
  mutate(stop = str_locate(string = sci_name, pattern = " ")[, 1],
         genus = substring(text = sci_name, first = 0, last = stop-1)) %>% 
  select(order, family, genus, sci_name, common_name, species, family, order)

# Load FAO country codes lookup table
FAO_cty_codes <- read.csv(file = here("raw_data","FAO","FAO_cty_codes.csv"), stringsAsFactors = F) %>% 
  select(country = UN_CODE, iso = Name_en)

# Load FAO aquaculture timeseries
FAO_ac_ts <- read.csv(file = here("raw_data","FAO","FAO_ac_ts.csv"), stringsAsFactors = F) %>%
  janitor::clean_names() %>% 
  filter(quantity > 0) %>% 
  left_join(FAO_sp_codes, by = "species") %>% 
  left_join(FAO_cty_codes, by = "country") %>% 
  janitor::clean_names() %>% 
  mutate(country = countrycode::countrycode(sourcevar = iso,
                                            origin = "iso3c",
                                            destination = "country.name",
                                            warn = F),
         source = "Aquaculture") %>% 
  group_by(year, production_area, country, iso, source, species, common_name, sci_name, genus, family, order) %>% 
  summarize(quantity = sum(quantity, na.rm = T), value = sum(value, na.rm = T)) %>% 
  ungroup()

# Load FAO fisheries production timeseries
FAO_fi_ts <- read.csv(file = here("raw_data","FAO","FAO_fi_ts.csv"), stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  filter(!unit == "no", quantity > 0) %>% 
  left_join(FAO_sp_codes, by = "species") %>% 
  left_join(FAO_cty_codes, by = "country") %>% 
  mutate(country = countrycode::countrycode(sourcevar = iso,
                                            origin = "iso3c",
                                            destination = "country.name",
                                            warn = F),
         source = "Catches",
         value = NA) %>% 
  select(year, production_area = fishing_area, country, iso, source, species, common_name, sci_name, genus, family, order, quantity, value)

# Join them together
FAO_ts <- rbind(FAO_ac_ts, FAO_fi_ts) %>% 
  arrange(year, country, source, species, common_name, sci_name, genus, family, order, quantity, value)


```

##Get salmon and lobster data
```{r}
## from FAO
fao_salmon_lobster <- FAO_ts %>% 
  mutate(lobster = grepl(pattern = "lobster", x = common_name),
         salmon = grepl(pattern = "salmon", x = common_name),
         general_name = case_when(lobster ~ "Lobster",
                                  TRUE ~ "Salmon")) %>% 
  filter(lobster | salmon) %>% 
  mutate(common_name = tolower(gsub(pattern = " nei", replacement = "", x = common_name)),
          common_name = gsub("\\s*\\([^\\)]+\\)", "", common_name)


```



#Code for GLM in R
```{r}



```



##Load RAM Legacy database
```{r}
ram <- read.csv(here("raw_data","RAM", "RAM.csv"), stringsAsFactors = F) %>% 
  janitor::clean_names()

```

```{r}
lobster_salmon_ram <- ram %>% 
  mutate(lobster = grepl(pattern = "lobster", x = stocklong),
         salmon = grepl(pattern = "salmon", x = stocklong),
         general_name = case_when(lobster ~ "Lobster",
                                  TRUE ~ "Salmon")) %>% 
  filter(lobster | salmon,
         tsid == "BdivBmsytouse-dimensionless")

```

###Graph of RAM Salmon and Lobster data
```{r}

lobster_salmon_ram %>% 
  group_by(tsyear, general_name) %>% 
  summarize(tsvalue = median(tsvalue, na.rm = T)) %>% 
  ggplot(aes(x = tsyear, y = tsvalue, color = general_name, group = general_name)) +
  geom_line(size = 1) +
  theme_bw()

```



