---
title: "FishBae"
subtitle: "EEMB 595 Final Project" 
author: "Phoebe Racine & Erin Winslow"
date: "5/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Task List for Phoebe & Erin
1. **Code GLM in STAN**
  - follow: https://datascienceplus.com/bayesian-regression-with-stan-part-1-normal-regression/
  - load data in STAN
  - fit data to model
  - Traceplot and posterior distribution
  - Pairwise correlation
  - Credible intervals around the parameters
  - Plot regression line
2. **Understand GLMs**
  - beta
  - read 3.3 of stats book
3. **combine Bay line w/ Juanca graph**
4.  **For Grace's analysis**
  - fill in outline
  - beautify graph
  - beautify regression table
5. **Presentation**
  - pull graphs from early AquaForce


#Report Outline
- **Introduction**
  - question
  - link to larger question
  - approach
  - data
  - literature
- **Approach**
  - log normal GLM - insert equation
    - normal prior
    - other equation considerations
    - for this model unlike Poisson, we don't need alpha or C
  - STAN
    - maybe explain basics of coding in STAN and connect cite resources
- **Analysis**
  - insert graphs
  - captions
- **Discussion**
  - difference between frequentist regression lines and Bayesian
- **Conclusion**
  - next steps
  

#Loading Packages
```{r, include=FALSE}
library(tidyverse)
library(stargazer)
library(ggplot2)
library(rstan)

# rstan_options(auto_write = TRUE)
# options(mc.cores = parallel::detectCores())

```


#Salmon and Lobster Code from Juanca
- Salmon-lobsert graph
- code for GLM in R
- code for GLM in JAGS

##Salmon-Lobster Graph
```{r}

simple_sam_lob <- fao_salmon_lobster  %>%
  group_by(year, source, general_name) %>% 
  summarize(quantity = sum(quantity)) %>% 
  filter(!(general_name == "Lobster" & source == "Aquaculture")) %>% 
  rename(Species = general_name)

  ggplot(simple_sam_lob, aes(x = year, y = quantity/1000000, color = source, linetype = Species, group = paste(source, Species))) +
  geom_line(size = 1) +
  scale_color_brewer(palette = "Set1") +
  geom_vline(xintercept = 1962, linetype = "dashed") +
  geom_text(aes(x = 1975, y = 2.5, label = "Salmon aquaculture begins"), color = "black", size = 4) +
  labs(x = "Year", y = "Quantities (million MT)", caption = "(Data from FAO: www.fao.org/fishery/statistics/)") +
  theme_minimal()

```

#STAN Log Normal GLM
- .stan file has been created for model
  - using a file allows you to use print statements and to dereference line numbers in error messages
  
**Preparing Data for STAN**
- this needs to all be numeric data
- need to create a matrix of predicted values
```{r}

binary_salmon_lobster <- fao_salmon_lobster %>%
  mutate(binary = ifelse(general_name == "Salmon", 1, 0)) %>% 
  select(year, source, quantity, binary) %>% 
  drop_na()
# View(binary_salmon_lobster)  

#Matrix for predictive values

# save predictions of the model in the new data frame 
# together with variable you want to plot against
# predicted_df <- data.frame(samlob_pred = predict(modeltestBAY, binary_salmon_lobster), hp=binary_salmon_lobster$quantity)
#View(predictedBAY_df)


```

*Example of data creation from Log Normal Example*
https://datascienceplus.com/bayesian-regression-with-stan-part-1-normal-regression/
```{r}
##Log Normal Example
#load libraries
library(rstan)
library(coda)
set.seed(20151204)
#the explanatory variables
dat<-data.frame(x1=runif(100,-2,2),x2=runif(100,-2,2))
#the model
X<-model.matrix(~x1*x2,dat)
#the regression slopes
betas<-runif(4,-1,1)
#the standard deviation for the simulated data
sigma<-1
#the simulated data
y_norm<-rnorm(100,X%*%betas,sigma)
#a matrix to get the predicted y values
new_X<-model.matrix(~x1*x2,expand.grid(x1=seq(min(dat$x1),max(dat$x1),length=20),x2=c(min(dat$x2),mean(dat$x2),max(dat$x2))))

View(new_X)

```


**Fitting Data for STAN Model**
- currently, there are errors due to data not properly formatted
```{r}
#From intro to STAN example
fit <- stan(file = 'FishBaeSTAN.stan', data = binary_salmon_lobster, 
            iter = 1000, chains = 4)


```


*Printing and Plotting STAN Model*
- pairs has error due to data
```{r}

print(fit)
plot(fit)
pairs(fit, pars = c("mu", "tau", "lp__"))

la <- extract(fit, permuted = TRUE) # return a list of arrays 
mu <- la$mu 

### return an array of three dimensions: iterations, chains, parameters 
a <- extract(fit, permuted = FALSE) 

### use S3 functions on stanfit objects
a2 <- as.array(fit)
m <- as.matrix(fit)
d <- as.data.frame(fit)


```


#Considering Approach
###What kind of GLM?
**GLM defintion**: In statistics, the generalized linear model (GLM) is a flexible generalization of ordinary linear regression that allows for response variables that have error distribution models other than a normal distribution. 
  - y = mx + b + e
  - e is an error term and can come from any distribution family
  - GLM is usually with a Poisson and a normal binominal for count data but you can also have it with beta or gamma distribution, multinomial or log normal

*Log-normal distribution*: a continuous probability distribution of a random variable whose logarithm is normally distributed. Thus, if the random variable X is log-normally distributed, then Y = ln(X) has a normal distribution. Likewise, if Y has a normal distribution, then the exponential function of Y, X = exp(Y), has a log-normal distribution. A random variable which is log-normally distributed takes only positive real values.

*Considerations for Log Normal Distribution*
Beta
- In Bayesian inference, the **beta distribution** is the conjugate prior probability distribution for the Bernoulli, binomial, negative binomial and geometric distributions. For example, the beta distribution can be used in Bayesian analysis to describe initial knowledge concerning probability of success such as the probability that a space vehicle will successfully complete a specified mission. The beta distribution is a suitable model for the random behavior of percentages and proportions.

**Poisson Generalized Linear Model**: a generalized linear model form of regression analysis used to model count data and contingency tables. Poisson regression assumes the response variable Y has a Poisson distribution, and assumes the logarithm of its expected value can be modeled by a linear combination of unknown parameters. A Poisson regression model is sometimes known as a log-linear model, especially when used to model contingency tables.
- We're treating quantities of aquaculture production and fisheries catch as count data.

**Poisson distribution**: a discrete probability distribution that expresses the probability of a given number of events occurring in a fixed interval of time or space if these events occur with a known constant rate and independently of the time since the last event (Frank A. Haight (1967). Handbook of the Poisson Distribution. New York: John Wiley & Sons).



#Resources
**We found a number of resources that are useful for learning how to use STAN.**
**To install**: https://github.com/stan-dev/rstan/wiki/Installing-RStan-on-Mac-or-Linux

Our own Dan Ovando gives us an incredibly helpful run down of how to use STAN: https://github.com/DanOvando/learn-stan/blob/master/documents/learn-stan.Rmd

**DBDA2Estan**, a collection of models from *Doing Bayesian Data Analysis* which are closely linked to previous examples made for JAGS. In most cases, the syntax is what's different.
https://github.com/jhoupt/DBDA2Estan

Examples of code from multiple Bayesian statistics package geared towards population ecology: https://github.com/stan-dev/example-models/wiki



